Safe Transferable Regions

### 2015-11-17 paper/type-system.tex
Edits.
Made "fresh" fix to the opsem.
Some edits to Sec.3.

\section{The Core Language}
\label{sec:type-system}

@@

We build on the Featherweight Generic Java (FGJ)~\cite{fgj} formalism
to formalize \name and its region type system. Our development reuses
notations from~\cite{fgj}, and relies on several of its definitions,
such as the definition of type well-formedness for the core
(region-free) language. We refrain from reproducing them here, and
instead encourage the reader to refer to~\cite{fgj,techrep}.

Fig~\ref{fig:fb-syntax} describes the syntax of our formal language,
which we call \fbname (\FB). Class types in \FB are region-annotated
variants of class types in FGJ (also called \emph{core types}). This
correspondence is reflected in the $T@\rgn$ syntax of a region type,
which is the simplest form of a region type describing an object of
core type $T$ contained in a region $\rgn$. We let $\rgn$ range over
static identifiers of regions in \FB. The only unboxed value in \FB is
$\unitval$ of type \unitZ. Rest are objects.

@@

To help state the type safety theorem, we extend \FB's expressions
with $\invalidexn$, and define the syntactic class of values:
\begin{smathpar}
\begin{array}{lclcl}
v & \in & \mathtt{values} & \coloneqq & \C{new}\; \fbN(\bar{v}) \\
\end{array}
\end{smathpar}
Type safety theorem is now stated thus:
\begin{theorem}
\emph{(\textbf{Type Safety})}
\label{thm:core-safety}
Let $\A_{\emptyset} = (\emptyset,\cdot,true)$ denote an empty context.
If $\hastyp{\A_{\emptyset}, \toprgn, \cdot}{e}{\tau}$, then either $e$
is a value, or $e \redsto \invalidexn$, or there exists an $e'$ such
that $e \redsto e'$ and $\hastyp{\A_{\emptyset}, \toprgn,
\cdot}{e'}{\tau}$.
\end{theorem}

>>>>

\section{\fbname}
\label{sec:type-system}

The purpose of \name's region type system is to enforce the key
invariant required for memory safety, namely that an object $o_1$ in a
region $R_1$ contains a reference to an object $o_2$ in $R_2$, only if
$R_2$ is guaranteed to outlive $R_1$. Intuitively, the invariant can
be enforced by (a). tracking \emph{outlives} relationships between
various regions in the program (b). tagging the C\# type of every
object in the program with its allocation region, and (c).  ensuring
that, when a reference is created, its target object ($o_2$) is
allocated in a region that is known to be in outlives relationship
with the object containing the reference. We now formally develop this
intuition via \fbname, our explicitly typed core language (with region
types) that incorporates the features introduced in the previous
section. \fbname builds on the Featherweight Generic Java
(FGJ)~\cite{fgj} formalism, and reuses notations and various
definitions from~\cite{fgj}, such as the definition of type
well-formedness for the core (region-free) language.

% In this section, we formally develop \fbname, a language that
% incorporates the features introduced in the previous section, as well
% as its type system, building on the Featherweight Generic Java
% (FGJ)~\cite{fgj} formalism.  Our development reuses notations and
% various definitions from~\cite{fgj}, such as the definition of type
% well-formedness for the core (region-free) language, which are
% elaborated further in the supplement~\cite{techrep}.

@@

Fig~\ref{fig:fb-syntax} describes the syntax of \fbname (\FB). Class
types in \FB are region-annotated variants of class types in FGJ (also
called \emph{core types}). This correspondence is reflected in the
$T@\rgn$ syntax of a region type, which is the simplest form of a
region type describing an object of core type $T$ contained in a
region $\rgn$. We let $\rgn$ range over static identifiers of regions
in \FB. The only unboxed value in \FB is $\unitval$ of type \unitZ.
Rest are objects.

@@

To help state the type safety theorem, we extend \FB's expressions
with $\invalidexn$, and define the syntactic class of values:
\begin{smathpar}
\begin{array}{lclcl}
v & \in & \mathtt{values} & \coloneqq & \C{new}\; \fbN(\bar{v}) \ALT
\lambdaexp{\ralloc}{\rhoalloc\rhobar\,|\,\phi}{\taubar \; \xbar}{e}\\
\end{array}
\end{smathpar}
Type safety theorem is now stated thus:
\begin{theorem}
\emph{(\textbf{Type Safety})}
\label{thm:core-safety}
Let $\A_{\emptyset} = (\emptyset,\cdot,true)$ denote an empty context.
If $\hastyp{\A_{\emptyset}, \toprgn, \cdot}{e}{\tau}$, then either $e$
is a value, or $e \redsto \invalidexn$, or there exists an $e'$ such
that $e \redsto e'$ and $\hastyp{\A_{\emptyset}, \toprgn,
\cdot}{e'}{\tau}$.
\end{theorem}
